Supplementary information: Variational theory of non-relativistic quantum electrodynamics

### 2019-01-29 supp2.tex
small changes supp

\textcolor{blue}{As discussed in the main text, the variational energies, as well as the energies from numerical diagonalization, are compared against the energies calculated from perturbation theory. We provide additional details on the perturbative calculations here. What we calculate is the energy shift of a state $|a,0\rangle$, where $a$ is any matter state, and $0$ denotes the photonic vacuum, according to the perturbation Hamiltonian $\frac{q}{m}Ap + \frac{q^2}{2m}A^2$, where $p$ is the momentum operator for the matter defined in the main text, and $A$ is the vector potential operator of the one-dimensional cavity in the absence of the matter. In particular, $A(z) = \sum_n\sqrt{\frac{\hbar}{\epsilon_0\omega_n L}}\sin\left(\frac{n\pi z}{L} \right) \left(a_n+a^{\dagger}_n\right)$, with $L$ the length of the cavity and $\omega_n = \frac{n\pi c}{L}$. Upon inspection, it is clear the the $A^2$ term leads to an energy shift at first-order in perturbation theory, while the $A\dot p$ term leads to an energy shift at second-order in perturbation theory. A straightforward application of perturbation theory leads to the following first-order energy shift from the $A^2$ term:
\begin{equation}
\delta E_a = \frac{q^2}{2m}\sum\limits_{n=1}^{N_p} \frac{\hbar}{\epsilon_0\omega_n L}\sin^2\left(\frac{n\pi z}{L}\right),
\end{equation}
where $z$ is the location of the matter in the cavity, and $N_p$ is the maximum photon number retained. Note that this value is dependent on the maximum photon number (although somewhat weakly) retained due to the logarithmically diverging nature of the sum, and so to compare variational, numerical, and perturbative calculations, the same number of modes are retained in each. A similarly straightforward application of perturbation theory leads to the energy shift coming from the $A\dot p$ term being given by:
\begin{equation}
\delta E_a = \frac{q^2}{m^2}\sum\limits_{b\neq a}\sum\limits_{n=1}^{N_p} \frac{\hbar}{\epsilon_0\omega_n L} \frac{|p_{ba}|^2\sin^2\left(\frac{n\pi z}{L} \right)}{\hbar(\omega_a - \omega_b - \omega_n)},
\end{equation}
where $b$ denotes a matter state different from $a$, and $p_{ba} = \langle b|p|a\rangle$.}

>>>>

\textcolor{blue}{As discussed in the main text, the variational energies, as well as the energies from numerical diagonalization, are compared against the energies calculated from perturbation theory. We provide additional details on the perturbative calculations here. What we calculate is the energy shift of a state $|a,0\rangle$, where $a$ is any matter state, and $0$ denotes the photonic vacuum, according to the perturbation Hamiltonian $\frac{q}{m}Ap + \frac{q^2}{2m}A^2$, where $p$ is the momentum operator for the matter defined in the main text, and $A$ is the vector potential operator of the one-dimensional cavity in the absence of the matter. In particular, $A(z) = \sum_n\sqrt{\frac{\hbar}{\epsilon_0\omega_n L}}\sin\left(\frac{n\pi z}{L} \right) \left(a_n+a^{\dagger}_n\right)$, with $L$ the length of the cavity and $\omega_n = \frac{n\pi c}{L}$. Upon inspection, it is clear that the $A^2$ term leads to an energy shift at first-order in perturbation theory, while the $A\cdot p$ term leads to an energy shift at second-order in perturbation theory. A straightforward application of perturbation theory leads to the following first-order energy shift from the $A^2$ term:
\begin{equation}
\delta E_a = \frac{q^2}{2m}\sum\limits_{n=1}^{N_p} \frac{\hbar}{\epsilon_0\omega_n L}\sin^2\left(\frac{n\pi z}{L}\right),
\end{equation}
where $z$ is the location of the matter in the cavity, and $N_p$ is the maximum photon number retained. Note that this value is dependent on the maximum photon number (although somewhat weakly) retained due to the logarithmically diverging nature of the sum, and so to compare variational, numerical, and perturbative calculations, the same number of modes are retained in each. A similarly straightforward application of perturbation theory leads to the energy shift coming from the $A\cdot p$ term being given by:
\begin{equation}
\delta E_a = \frac{q^2}{m^2}\sum\limits_{b\neq a}\sum\limits_{n=1}^{N_p} \frac{\hbar}{\epsilon_0\omega_n L} \frac{|p_{ba}|^2\sin^2\left(\frac{n\pi z}{L} \right)}{\hbar(\omega_a - \omega_b - \omega_n)},
\end{equation}
where $b$ denotes a matter state different from $a$, and $p_{ba} = \langle b|p|a\rangle$.}
